---
title: "Homework 3"
author: "Alex Wako"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(astsa)
library("readxl")
```

1

```{r 1}
x1 = UnempRate
logx1 = log(x1); dlogx1 = diff(logx1); ddlogx1 = diff(dlogx1, 12)
plot.ts(cbind(x1, logx1, dlogx1, ddlogx1), main = "")
sarima(ddlogx1, 0, 1, 1, 1, 0, 0, 12) # The Seasonal ARIMA model
prediction1 <- sarima.for(x1, 12, 0, 1, 1, 1, 0, 0, 12) # Forecast of the next 12 months
prediction1$pred
```
# 2

```{r 2}
temp <- read.table("Problem2.txt", header = T, sep = ",")
x2 <- ts(temp$Value, start = c(1948, 1), frequency = 12)
logx2 = log(x2); dlogx2 = diff(logx2); ddlogx2 = diff(dlogx2, 12)
plot.ts(cbind(x2, logx2, dlogx2, ddlogx2), main = "")
sarima(ddlogx2, 0, 1, 1, 1, 0, 0, 12) # The Seasonal ARIMA model
prediction2 <- sarima.for(x2, 12, 1, 0, 1, 1, 0, 0, 12) # Forecast of the next 12 months
prediction2$pred
```
\
Comparing the two plots created by the Seasonal ARIMA(0, 1, 1) * (1, 0, 0), the forecast found in the first plot stays relatively similar to the actual data found in the second plot. A large number of unemployment occurred during the first quarter of 2020, likely due to the rise of COVID-19, and businesses needing to cut down on expenses, but by 2022, the time series plot went back down to follow the trend of the time series plot. From the forecast created from the second plot, a likely prediction is a small rising trend in unemployment for the next 12 months with the confidence intervals showing that to be more than likely that an even or rising trend will occur.
